---
output:
  pdf_document: default
  html_document: default
---
install.packages(Rmisc)
install.packages("data.table")
install.packages("stringr")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("readxl")
options(scipen = 999)
---
title: "LabPhon_Coding"
author: "Myeongju"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(Rmisc))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(readxl))
options(scipen=999)
```

## Preprocessing

# Defining a new function
This enables us to subset later in a way such that we only keep the data which doesn't include the specified values.

```{r}
`%notin%` <- Negate(`%in%`)
```

# Data

```{r pressure, echo=FALSE}
df <- read_excel("Results_F0_Data.xlsx")
names(df)
```

```{r}
summary(df)
```

```{r}
plot(density(df$`Mean(Hz)`, na.rm=TRUE))
```
# Separating the data set into DV subsets
```{r}
df_f0 <- df[, c("Year", "Country", "Mean(Hz)","StDev(Hz)")]
```

```{r}
f0_averaged <- summarySE(df_f0, measurevar = "Mean(Hz)", groupvars = c("Year", "Country"), na.rm=T)
f0_averaged <-f0_averaged[,c(1,2,4)]
colnames(f0_averaged)[3] <- "avg_f0"
newdata <- f0_averaged[-c(73), ]
```

```{r}
summary(f0_averaged)
summary(newdata)
```

```{r}
write.csv(newdata, "F0_averaged.csv")
```


